Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2020-05-06 article.tex
Evolution method describe for two-state Markov chain

To detect evolution, we determine the probability 
that all of the observed TMH mutations have arisen by chance,
with regards to being detected/presented by the immune system.
If this probability is lower than 5\%, we will state 
that this unlikely situation is most likely caused by natural selection.

\begin{figure}
  \centering
  \input{markov_chain.latex}
  \caption{
    State transitions in TMHs caused by mutations, with respect to the
    detection by the immune system. Any TMH is either in a detected, $D$, or
    undetected, $U$, state. Mutations cause state transitions with probability
    $p$. For example, $p_{DU}$ denotes the probability that a detected
    TMH mutations into an undetected one.
  }
  \label{fig:markov_chain}
\end{figure}

Transition matrix from \cite{dang2010flu}.

>>>>

To detect evolution, we determine the probability 
that all of the observed TMH mutations have arisen by chance,
with regards to being detected/presented by the immune system.
To do so, we calculate the probability of observing the complete
dataset, for which we need to calculate the probability 
of each individual mutation.

\paragraph{System depicted as a Markov chain}

To calculate the probability all of our observations happened by chance,
we view our experiment as a two-state Markov chain, as depicted in
figure \ref{fig:markov_chain}. The two nodes depict the
two possible detection state a polypeptide is in, which is
either undetected $U$, or detected $D$ by the immune system.
The edges depict the probability of a state transition by mutation,
for example, $p_{DU}$ denotes the probability that a detected
polypeptide mutates into an undetected one.
We focus on polypeptides that are TMH and we assume a mutation that 
renders the TMH non-functional as lethal.

\begin{figure}
  \centering
  \input{markov_chain.latex}
  \caption{
    State transitions in TMHs caused by mutations, with respect to the
    detection by the immune system. Any TMH is either in a detected, $D$, or
    undetected, $U$, state. Mutations cause state transitions with probability
    $p$. For example, $p_{DU}$ denotes the probability that a detected
    TMH mutates into an undetected one.
  }
  \label{fig:markov_chain}
\end{figure}

\paragraph{Statistics}

When we have calculated the probability of each transition (as described below),
we can simply count the number of observed transitions.
Because the probability of transitions differ per polypeptide,
the total system has the properties of a Poisson binomial distribution,
where, for example, the expected number of transitions is the sum of individual
probabilities. We calculate the probability of the observing our
number of transitions.
If this probability is lower than 5\%, we will state 
that this unlikely situation is most likely caused by natural selection.

\paragraph{Calculating one polypeptide's probability}

We calculate the probability of a transition per polypeptide.
For any polypeptide, we can generate all sequences that differ
in only one amino acids. Of these sequences, we only keep the ones
that result in a TMH. Of each of the remaining TMH sequences, we predict
which mutants are detected by the immune system. Because not all 
AA mutations are equally likely, we correct for the AA transition
rates using the FLU transition matrix (\cite{dang2010flu}),
which is a transition matrix derived from observed influenza AA mutations.

\paragraph{Assumption that epitopes are immunogenic}
